SMT-Boosted Security Types for Low-Level MPC

### 2025-11-13 toplas26/metaty.tex
checkpoint

We also introduce dependent $\Pi$ types for functions, that allow
dependence on $\stringty$ and $\cidty$ types.
\begin{definition}[$\metaprot$ Function Types]
  Dependent function types are of the syntactic form
  $\Pi y_1,\ldots,y_n. \notg{\Gamma} \rightarrow (\notg{\Gamma}, \notg{R})$.
  where $y_1,\ldots,y_n$ range over $\stringty$ or $\cidty$ types.
\end{definition}
As defined in \TirName{Fn}, a dependent function type $\Pi
y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2},
\notg{R})$ specifies that, given share typing assumptions in
$\notg{\Gamma_1}$, application the function results in the bindings in
$\notg{\Gamma_2}$ by using the random values in $\notg{R}$. It is
always the case that the type is closed, i.e., only $y_1,\ldots,y_n$
can occur free in $\notg{\Gamma_1}$, $\notg{\Gamma_2}$ $\notg{R}$.
Type checking for functions is accomplished by using fresh ``witness''
values of $\stringty$ and $\cidty$ types to instantiate dependent
variables, similarly to the \TirName{GenEntails} rule. As defined in
the \TirName{App} rule, typing ensures that application of the
function instantiates the dependencies, checks that the instantiation
of $\notg{\Gamma_1}$ is contained in the share typing environment, and
adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
the instantiation of $\notg{R}$ with linearity enforced.

@@

Returning to the GMW \ttt{and} gate example introduced in Figure \ref{fig-gmwprepost},
we can extend the definition of \ttt{andgmw} with type annotations to ensure a
valid share typing, as in Figure \ref{fig-gmwsharetypes}. Here a coercion
is needed to specify that the gate output are shares obtained using
$\ttt{r[z]@1}$ as an OTP. The type for $\ttt{andgmw}$ can then be
obtained as:
$$
\ttt{andgmw} : \Pi x,y,z . \varnothing \rightarrow (\setit{[m[z]@1,m[z]@2] : \sharety}, \setit{r[z]@1})  
$$
Note that the function does not require any share type preconditions, since
the (sub)protocol it defines interactively generates new shares using the OTP.

\begin{fpfig}[t]{GMW $\ttt{and}$ gate with share type annotations.}{fig-gmwsharetypes}
\begin{verbatimtab}
andgmw(z:string, x:string, y:string) {
  let table = andtablegmw(x,y,z) in
  m[z]@2 := OT4((m[x], m[y])@2, table.row4, table.row3, table.row2, table.row1)@1;
  m[z]@1 := r[z]@1;
  m[z]@2 as (RECON(x) * RECON(y)) + r[z]@1;
  [m[z]@1,m[z]@2] : shares
}
postcondition:  ( RECON(z) == RECON(x) * RECON(y))
\end{verbatimtab}
\end{fpfig}

>>>>

We also introduce dependent $\Pi$ types for functions, that allow
dependence on $\stringty$ and $\cidty$ types.
\begin{definition}[$\metaprot$ Function Types]
  Dependent function types are of the syntactic form
  $\Pi y_1,\ldots,y_n. \notg{\Gamma} \rightarrow (\notg{\Gamma}, \notg{R})$.
  where $y_1,\ldots,y_n$ range over $\stringty$ or $\cidty$ types.
\end{definition}
As defined in \TirName{Fn}, a dependent function type $\Pi
y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2},
\notg{R})$ specifies that, given share typing assumptions in
$\notg{\Gamma_1}$, application the function results in the bindings in
$\notg{\Gamma_2}$ by using the random values in $\notg{R}$. It is
always the case that the type is closed, i.e., only $y_1,\ldots,y_n$
can occur free in $\notg{\Gamma_1}$, $\notg{\Gamma_2}$ $\notg{R}$.
Type checking for functions is accomplished by using fresh ``witness''
values of $\stringty$ and $\cidty$ types to instantiate dependent
variables, similarly to the \TirName{GenEntails} rule. As defined in
the \TirName{App} rule, typing ensures that application of the
function instantiates the dependencies, checks that the instantiation
of $\notg{\Gamma_1}$ is contained in the share typing environment, and
adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
the instantiation of $\notg{R}$ with linearity enforced. A practical
programming detail is where the share type environment precondition
$\notg{\Gamma_1}$ is defined.  In our implementation, we write these
as an additional component of \ttt{precondition} annotations, as
illustrated in extended examples discussed in Section
\ref{section-examples}.

@@

Returning to the GMW \ttt{and} gate example introduced in Figure \ref{fig-gmwprepost},
we can extend the definition of \ttt{andgmw} with type annotations to ensure a
valid share typing, as in Figure \ref{fig-gmwsharetypes}. Here a coercion
is needed to specify that the gate output are shares obtained using
$\ttt{r[z]@1}$ as an OTP. The type for $\ttt{andgmw}$ can then be
obtained as:
$$
\ttt{andgmw} : \Pi x,y,z . \varnothing \rightarrow (\setit{[m[z]@1,m[z]@2] : \sharety}, \setit{r[z]@1})  
$$
Note that the function does not require any share type preconditions, since
the (sub)protocol it defines interactively generates new shares using an OTP
($r[z]@1$).

\begin{fpfig}[t]{GMW $\ttt{and}$ gate with share type annotations.}{fig-gmwsharetypes}
\begin{verbatimtab}
andgmw(z:string, x:string, y:string) {
  let table = andtablegmw(x,y,z) in
  m[z]@2 := OT4((m[x], m[y])@2, table.row4, table.row3, table.row2, table.row1)@1;
  m[z]@1 := r[z]@1;
  m[z]@2 as (RECON(x) * RECON(y)) + r[z]@1;
  [m[z]@1,m[z]@2] : shares
}
postcondition: (RECON(z) == RECON(x) * RECON(y))
\end{verbatimtab}
\end{fpfig}

\subsubsection{Soundness of $\metaprot$ Type Checking}

\begin{lemma}
  Given $\prog_1,\Gamma_1,R_1 : {OK}$ and $\cmd \redx \prog_2$ and $\htrip{\eqs_1}{\cmd}{\eqs_2}$ with
  $\eqs_1$ satisfiable and $\toeq{\prog_1} \models \eqs_1 \wedge \toeq{\xdefs_1}$ and
  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$.
  Then $\prog_1;\prog_2,\Gamma_2,R_2 : {OK}$ and $\toeq{\prog_1;\prog_2} \models \toeq{\xdefs_2}$.
\end{lemma}

\begin{theorem}
  Given $\htrip{\eqspre}{\ttt{main()}}{\eqs}$ and $\ttt{main} : \Gamma_1 \rightarrow (\Gamma_2,R)$
  and $\main() \redx \prog$ and $\progpre,\Gamma_1,R' : {OK}$ with $R' \cap R = \varnothing$
  and $\toeq{\progpre} \models \eqspre$. Then $(\progpre;\prog),\Gamma_2,R : {OK}$.
\end{theorem}

% DEPRECATED
\begin{comment}
\begin{lemma}
  Assume $\cmd \redx \prog$ and $\htrip{\eqs_1 \wedge \toeq{\xdefs_1}}{\prog}{\eqs_2}$ and
  $\shtrip{\Gamma_1,R_1,\xdefs_1}{\cmd}{\Gamma_2,R_2,\xdefs_2}$ and   
  $\eqs_1 \wedge \toeq{\xdefs_1},\clean(\Gamma_1),R_1 : {OK}$.
  Then $\eqs_1 \wedge \toeq{\xdefs_1} \wedge \toeq{\prog} \vdash \clean(\Gamma_1),R_1 \stredx^* \clean(\Gamma_2),R_2$.
\end{lemma}

### 2025-11-13 toplas26/sty.tex
checkpoint

We write $\stredx^*$ to denote the reflexive, transitive
closure of $\stredx$. For any protocol $\prog$, the goal of the
type analysis is to show that all interactive messages can be classified--
that is, to show that $\eqspre \wedge \toeq{\prog} \vdash
\varnothing,\varnothing \stredx^* \Gamma,R$ where all
interactive messages are classified in $\Gamma$.

@@

Now we can prove our share typing soundness result, showing that share
typing enforces noninteference modulo output, with our main result
in Theorem \ref{theorem-sty}. We begin by defining a key well-formedness
property of typing components that we show is preserved by non-interactive
typing judgements.
\begin{definition}
  We write $\eqs,\Gamma,R : {OK}$ iff both of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $([\phi_1,\phi_2] : \sharety) \in \Gamma$,
    we have $\eqs \models \phi_1 \eop \phi - \rx{x}{\cid'} \wedge
    \phi_2 \eop \rx{y}{\cid'}$ for
    some $\phi$ and unique $\rx{x}{\cid'} \in R$, with $\cid \ne \cid'$.
  \item For all $(\pubx{\phi} : \pubty) \in \Gamma$,
    we have $\eqs \models \phi \eop \phi' - \rx{x}{\oid}$ for
    some $\phi'$ and unique $\rx{x}{\oid} \in R$.
  \end{enumerate}
\end{definition}

@@

Based on the preceding Lemma, we can show that noninterence with respect
to classified interactive variables is maintained by typing judgements. 
\begin{lemma}
  \label{lemma-styinteractive}
  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}$  and
  $\iov(\prog) = S \cup V \cup O$
  and $\eqspre \wedge \toeq{\prog} \vdash \Gamma_1,R_1 \stredx \Gamma_2,R_2$.
  Letting $X_1 \defeq \ov(\Gamma_1) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
  for $\cid_1 \ne \cid_2$ assume also:
  $$
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_1 \cup O}
  = 
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
  $$
  Then, letting $X_2 \defeq \ov(\Gamma_2) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
  we have:
  $$
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_2 \cup O}
  = 
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
  $$
\end{lemma}
Now we can prove our main Theorem, which follows immediately 
by Lemma \ref{lemma-styinteractive} and induction on $|\Gamma|$
(i.e., the number of bindings in the environment $\Gamma$ obtained
by a complete typing). 
\begin{theorem}[Share Type Soundness]
  \label{theorem-sty}
  Assume given $\prog$ with 
  $\iov(\prog) = S \cup V \cup O$
  and $\eqspre \wedge \toeq{\prog} \vdash \varnothing,\varnothing \stredx^*
  \Gamma,R$ where $\realviews{\setit{\cid_1}}{\setit{\cid_2}} \cup O
  \subseteq \ov(\Gamma)$ for all $\cid_1$, $\cid_2$ with $\cid_1 \ne \cid_2$.
  Then $\prog$ satisfies noninterference modulo output.
\end{theorem}

>>>>

We write $\stredx^*$ to denote the transitive
closure of $\stredx$. For any protocol $\prog$, the goal of the
type analysis is to show that all interactive messages can be classified--
that is, to show that $\eqspre \wedge \toeq{\prog} \vdash
\varnothing,\varnothing \stredx^* \Gamma,R$ where all
interactive messages are classified in $\Gamma$.

@@

Now we can prove our share typing soundness result, showing that share
typing enforces noninteference modulo output, with our main result
in Theorem \ref{theorem-sty}. We begin by defining a key well-formedness
property of typing components that we show is preserved by non-interactive
typing judgements.

\begin{definition}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $\toeq{\prog} \models \phi_i^a \eop \rx{w}{\cid_a} \wedge \phi_i^b \eop \phi - \rx{w}{\cid_a}$ for
    some $\phi$ and $\rx{w}{\cid_a}$ then $\rx{w}{\cid_a} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}

\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
    \end{enumerate}
\end{lemma}

\begin{lemma}
  Given $\prog,\Gamma_1,R_1  : {OK}$
  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
\end{lemma}

% DEPRECATED
\begin{comment}
\begin{definition}
  We write $\eqs,\Gamma,R : {OK}$ iff both of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $([\phi_1,\phi_2] : \sharety) \in \Gamma$,
    we have $\eqs \models \phi_1 \eop \phi - \rx{x}{\cid'} \wedge
    \phi_2 \eop \rx{y}{\cid'}$ for
    some $\phi$ and unique $\rx{x}{\cid'} \in R$, with $\cid \ne \cid'$.
  \item For all $(\pubx{\phi} : \pubty) \in \Gamma$,
    we have $\eqs \models \phi \eop \phi' - \rx{x}{\oid}$ for
    some $\phi'$ and unique $\rx{x}{\oid} \in R$.
  \end{enumerate}
\end{definition}

@@

Based on the preceding Lemma, we can show that noninterence with respect
to classified interactive variables is maintained by typing judgements. 
\begin{lemma}
  \label{lemma-styinteractive}
  Assume given $\eqspre \wedge \toeq{\prog}, \Gamma_1,R_1: {OK}$  and
  $\iov(\prog) = S \cup V \cup O$
  and $\eqspre \wedge \toeq{\prog} \vdash \Gamma_1,R_1 \stredx \Gamma_2,R_2$.
  Letting $X_1 \defeq \ov(\Gamma_1) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
  for $\cid_1 \ne \cid_2$ assume also:
  $$
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_1 \cup O}
  = 
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
  $$
  Then, letting $X_2 \defeq \ov(\Gamma_2) \cap \realviews{\setit{\cid_1}}{\setit{\cid_2}}$,
  we have:
  $$
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}}  \cup X_2 \cup O}
  = 
  \condd{\progtt(\prog)}{S_{\setit{\cid_1}}}{S_{\setit{\cid_2}} \cup O}
  $$
\end{lemma}
Now we can prove our main Theorem, which follows immediately 
by Lemma \ref{lemma-styinteractive} and induction on $|\Gamma|$
(i.e., the number of bindings in the environment $\Gamma$ obtained
by a complete typing). 
\begin{theorem}[Share Type Soundness]
  \label{theorem-sty}
  Assume given $\prog$ with 
  $\iov(\prog) = S \cup V \cup O$
  and $\eqspre \wedge \toeq{\prog} \vdash \varnothing,\varnothing \stredx^*
  \Gamma,R$ where $\realviews{\setit{\cid_1}}{\setit{\cid_2}} \cup O
  \subseteq \ov(\Gamma)$ for all $\cid_1$, $\cid_2$ with $\cid_1 \ne \cid_2$.
  Then $\prog$ satisfies noninterference modulo output.
\end{theorem}
\end{comment}
